---
title: "Bellabeat FitBit"
author: "Roshini Priya"
date: "2023-02-27"
output: html_document
---

##Loading Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyr)
```
##Importing datasets
```{r echo=TRUE}
activity<-read.csv("dailyActivity_merged.csv")
intensities<-read.csv("hourlyIntensities_merged.csv")
calories<-read.csv("hourlyCalories_merged.csv")
sleep<-read.csv("sleepDay_merged.csv")
weight<-read.csv("weightLogInfo_merged.csv")
```
##Formatting

```{r echo=TRUE}
# intensities
intensities$ActivityHour=as.POSIXct(intensities$ActivityHour, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
intensities$time <- format(intensities$ActivityHour, format = "%H:%M:%S")
intensities$date <- format(intensities$ActivityHour, format = "%m/%d/%y")
# calories
calories$ActivityHour=as.POSIXct(calories$ActivityHour, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
calories$time <- format(calories$ActivityHour, format = "%H:%M:%S")
calories$date <- format(calories$ActivityHour, format = "%m/%d/%y")
# activity
activity$ActivityDate=as.POSIXct(activity$ActivityDate, format="%m/%d/%Y", tz=Sys.timezone())
activity$date <- format(activity$ActivityDate, format = "%m/%d/%y")
# sleep
sleep$SleepDay=as.POSIXct(sleep$SleepDay, format="%m/%d/%Y %I:%M:%S %p", tz=Sys.timezone())
sleep$date <- format(sleep$SleepDay, format = "%m/%d/%y")
```
##Data Summary

```{r echo=TRUE}
n_distinct(activity$Id)
n_distinct(calories$Id)
n_distinct(intensities$Id)
n_distinct(sleep$Id)
n_distinct(weight$Id)
```
##Data Analysing

```{r echo=TRUE}
# activity
activity %>%  
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes, Calories) %>%
  summary()

# explore num of active minutes per category
activity %>%
  select(VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes) %>%
  summary()

# calories
calories %>%
  select(Calories) %>%
  summary()
# sleep
sleep %>%
  select(TotalSleepRecords, TotalMinutesAsleep, TotalTimeInBed) %>%
  summary()
# weight
weight %>%
  select(WeightKg, BMI) %>%
  summary()
```
## Merging of data

```{r echo=TRUE}
merged_data <- merge(sleep, activity, by=c('Id', 'date'))
head(merged_data)
```
## Including Plots for Visualization

1. A positive correlation is seen between the total steps and the calories burnt by the users as follows

```{r echo=TRUE}
ggplot(data=activity, aes(x=TotalSteps, y=Calories)) + 
  geom_point() + geom_smooth() + labs(title="Total Steps vs. Calories")
```
2. Active hours - 5pm to 7pm which can be seen below

```{r echo=TRUE}
int_new <- intensities %>%
  group_by(time) %>%
  drop_na() %>%
  summarise(mean_total_int = mean(TotalIntensity))

ggplot(data=int_new, aes(x=time, y=mean_total_int)) + geom_histogram(stat = "identity", fill='darkblue') +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title="Average Total Intensity vs. Time")
```
3. A negative correlation is seen between Sedentary time and Sleep time as follows

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(data=merged_data, aes(x = TotalMinutesAsleep, y = SedentaryMinutes)) + geom_point(color='darkblue')+geom_smooth()+ labs(title="Minutes Asleep vs. Sedentary Minutes")
```


##KEY INSIGHTS & SUGGESTIONS:

1. Since from the summary it is seen that the Average Total Steps by the users are less than 8000, a notification can be sent to the users to encourage them to walk more stating its benefits for a good health.
2. Since the most activity is happening between 5pm-7pm, a notification to remind the users to go for a walk or run will be useful.
3. Average Sedentary time is more which is unhealthy, so a notification to stay active and reduce the sedentary time will also be helpful to monitor their activities.
4. Fitbit does not have information about user hydration level, so bellabeat can try to incorporate this information as well since staying hydrated is also a key factor for a healthy lifestyle.
